Improving Software Maintainability through Automated Refactoring of Code Clones

### 2019-08-15 icse/paper.tex
Refactoring

\begin{abstract}
Duplication in source code is often seen as one of the most harmful types of technical debt, because it increases the size of the codebase and creates implicit dependencies between fragments of code. To remove such anti-patterns, a developer should refactor the codebase. There are many tools that assist in this process. The thresholds and prioritization of the clones detected by such tools can be improved by considering the maintainability of the codebase after a detected clone would be refactored.

@@

\subsubsection{Characteristics of the extracted method}
We define the following characteristics of the extracted method and/or the call:
\begin{itemize}
\item \textbf{Statements}: The number of statements in the body of the method.
\item \textbf{Tokens}: The number of tokens in the body of the method.
\item \textbf{Relation}: The relation category (Sec.~\ref{sec:relation}) by which this methods' location was determined.
\item \textbf{Returns}: Whether the method calls return, declare, assign or don't use any data from the extracted method.
\item \textbf{Parameters}: The number of parameters the extracted method has.
\end{itemize}
We hypothesize that these characteristics are the main factors influencing the impact on the maintainability of the system as a result of refactoring the clone.

@@

\subsection{Calculating a maintainability score}
In this study we use four metrics to determine maintainability \ref{sec:metrics}. For our experiments, we want to aggregate the scores obtained by these metrics to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability.
  \item The obtained increase of the metric divided by the average distance from zero over our corpus weights the metric equally against the other metrics.
\end{itemize}
We calculate the average deviation as follows:
\begin{equation}\label{eq:scoredev}
A = \frac{\sum_{x \in M} abs(x)}{|M|}
\end{equation}
Where \textit{M} is the multiset of all change in a certain metric for all refactorings over all systems in  obtained for a certain metric over our corpus and \textit{|M|} is the cardinality of this set. We then calculate the maintainability for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
A_M = \frac{dup}{A_{dup}} + \frac{com}{A_{com}} + \frac{par}{A_{par}} + \frac{vol}{A_{vol}}
\end{equation}
Where \textit{dup} is the decrease in duplication, \textit{com} is the decrease in complexity, \textit{par} is the decrease is method parameters and \textit{vol} is the decrease in system volume after the refactoring is applied. When comparing the maintainability of a set of refactorings we divide the sum of the maintainability score by the amount of clone instances of all refactored clones in the set.

@@

\subsection{Clone refactorability}
%To what extent can found clones be refactored through method extraction, without requiring additional transformations.
Table \ref{tab:refactorability} shows to what extend clone classes can be refactored by using the ``Extract Method'' refactoring technique. The second column shows our measurements for the complete systems (just like the former experiments). The third column shows our measurements when restricting our search to method bodies. The amount that can be extracted increases because, when restricting our search to method bodies, we do not exclude declarations that can obstruct the possibility of method extraction (for instance a cloned method signature).

\begin{table}[]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textit{\textbf{Category}} & \textit{\textbf{All}} & \textit{\textbf{Method Body}} \\ \midrule
Can Be Extracted & 24,157 & 26,109 \\
Is Not A Partial Method & 21,625 & 0 \\
Top-level Node is not a Statement & 19,887 & 4,607 \\
Spans Part of a Block & 12,964 & 13,460 \\
Multiple Return Values & 5,622 & 6,131 \\
Complex Control Flow & 1,106 & 1,216 \\ \bottomrule
\end{tabular}
\caption{Number of clones that can be extracted using the ``Extract Method'' refactoring technique}
\label{tab:refactorability}
\end{table}s

\subsection{Thresholds}
I think the ultimate goal with this thesis is to do experiments with different clone thresholds. Which thresholds give clones that we should refactor? For this, we will measure the maintainability of the refactored source code over different thresholds. These thresholds range from minimum clone size, variability and gap size.

>>>>

\begin{abstract}
Duplication in source code is often seen as one of the most harmful types of technical debt, because it increases the size of the codebase and creates implicit dependencies between fragments of code. To remove such anti-patterns, a developer should refactor the codebase. There are many tools that assist in this process. The thresholds and prioritization of the clones detected by such tools can be improved by considering the maintainability of the codebase after a detected clone would be refactored. We perform an exploratory study into the data that can be gathered from comparing before- and after snapshots of an automatically refactored system.

@@

\subsubsection{Characteristics of the extracted method}\label{sec:characteristics}
We define the following characteristics of the extracted method and/or the call:
\begin{itemize}
\item \textbf{Statements}: The number of statements in the body of the method.
\item \textbf{Tokens}: The number of tokens in the body of the method.
\item \textbf{Relation}: The relation category (Sec.~\ref{sec:relation}) by which this methods' location was determined.
\item \textbf{Returns}: Whether the method calls return, declare, assign or don't use any data from the extracted method.
\item \textbf{Parameters}: The number of parameters the extracted method has.
\end{itemize}
We hypothesize that these characteristics are the main factors influencing the impact on the maintainability of the system as a result of refactoring the clone.

@@

\subsection{Calculating a maintainability score}\label{sec:metricformula}
In this study we use four metrics to determine maintainability \ref{sec:metrics}. For our experiments, we want to aggregate the scores obtained by these metrics to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability.
  \item The obtained increase of the metric divided by the average distance from zero over our corpus weights the metric equally against the other metrics.
\end{itemize}
We calculate the average deviation as follows:
\begin{equation}\label{eq:scoredev}
A = \frac{\sum_{x \in M} abs(x)}{|M|}
\end{equation}
Where \textit{M} is the multiset of all change in a certain metric for all refactorings over all systems in  obtained for a certain metric over our corpus and \textit{|M|} is the cardinality of this set. We then calculate the maintainability for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
A_M = \frac{dup}{A_{dup}} + \frac{com}{A_{com}} + \frac{par}{A_{par}} + \frac{vol}{A_{vol}}
\end{equation}
Where \textit{dup} is the decrease in duplication, \textit{com} is the decrease in complexity, \textit{par} is the decrease is method parameters and \textit{vol} is the decrease in system volume after the refactoring is applied. When comparing the maintainability of a set of refactorings we divide the sum of the maintainability score by the amount of clone instances of all refactored clones in the set.

@@

\subsection{Extract Method}
%To what extent can found clones be refactored through method extraction, without requiring additional transformations.
Table \ref{tab:refactorability} shows to what extend clone classes can be refactored by using the ``Extract Method'' refactoring technique. The second column shows our measurements for the complete systems (just like the former experiments). The third column shows our measurements when restricting our search to method bodies. The amount that can be extracted increases because, when restricting our search to method bodies, we do not exclude declarations that can obstruct the possibility of method extraction (for instance a cloned method signature).

\begin{table}[]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
\textit{\textbf{Category}} & \textit{\textbf{All}} & \textit{\textbf{Method Body}} \\ \midrule
Can Be Extracted & 24,157 & 26,109 \\
Is Not A Partial Method & 21,625 & 0 \\
Top-level Node is not a Statement & 19,887 & 4,607 \\
Spans Part of a Block & 12,964 & 13,460 \\
Multiple Return Values & 5,622 & 6,131 \\
Complex Control Flow & 1,106 & 1,216 \\ \bottomrule
\end{tabular}
\caption{Number of clones that can be extracted using the ``Extract Method'' refactoring technique}
\label{tab:refactorability}
\end{table}

\subsection{Refactoring}
%I think the ultimate goal with this thesis is to do experiments with different clone thresholds. Which thresholds give clones that we should refactor? For this, we will measure the maintainability of the refactored source code over different thresholds. These thresholds range from minimum clone size, variability and gap size.
In our corpus CloneRefactor has refactored 12.710 clone classes and measured the change in indicated metrics (see Sec.~\ref{sec:metrics}). Using the presented formulas (see Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring.

\subsubsection{Statements}
